---
title: "Data Visualization with Highcharter in R"
author: "Nishant"
date: "June 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(highcharter)
library(ggplot2)
library(plotly)
library(dplyr)
data("pokemon")
```


# Data Visualization with Highcharter in R

## Learn how to use Highcharter to create visualization that creates a buzz

I think data visualization is the best technique to show any descriptive and analytics reporting on a chunk of data.
I am a type of person who loves data visualization, you can show the whole story in single screen well that also depends on the data complexity. 
If you are reading this tutorial then I think you must be aware of `Ggplot2` package in R which is used to generate some awesome charts for analysis but somehow lacks dynamic properties, well moving forward came `Plotly` which provided an extension to present dynamic charts, It can also be used as wrapper of ggplot2 object and making a dynamic chart using ggplotly function as shown in the example.

```{r PlotlyExample, message=FALSE, warning=FALSE}



```

But Its a separate topic may be in future you can get a tutorial over it.


Coming Back to Highcharter, So it is a `R` wrapper of **HighCharts** javascript library and its module.

The main features of this packege is:

* You can create various charts with same style like scatter, bubble, time series, heatmaps, treemaps, bar charts etc..
* It supports various R objects.
* It supports Highstocks Charts, Choloropleths.
* It do have piping style which is loved by all R users and programmer.
* A large variety of themes with awesome looks.

Lets get down to the business and create some visulaization with Highcharter following the features mentioned above:

## Creating some Basic Charts

Lets choose a dataset, I am going to take Pokemon dataset also provided in Highcharter package.
Have a glimpse of the datset.

```{r, message=FALSE, warning=FALSE}

glimpse(pokemon)

```

Lets plot a bar chart.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "bar", hcaes(x = type_1, y = n))

```


So you got a bar chart with repect of of type 1 category of the pokemons.

Suppose you want a column chart instead then the only variable that you need to change is the type to column.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "column", hcaes(x = type_1, y = n))


```

Lets plot some other charts too.

### Treemap
```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "treemap", hcaes(x = type_1, value = n, color = n))


```

### 
