---
title: "Data Visualization with Highcharter in R"
author: "Nishant"
date: "June 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(highcharter)
library(ggplot2)
library(dplyr)
library(broom)
data("pokemon")
```


# Data Visualization with Highcharter in R

## Learn how to use Highcharter to create visualization that creates a buzz

I think data visualization is the best technique to show any descriptive and analytics reporting on a chunk of data.
I am a type of person who loves data visualization, you can show the whole story in single screen well that also depends on the data complexity. 
If you are reading this tutorial then I think you must be aware of `Ggplot2` package in R which is used to generate some awesome charts for analysis but somehow lacks dynamic properties.


Coming Back to Highcharter, So it is a `R` wrapper of **HighCharts** javascript library and its module.

The main features of this package are:

* You can create various charts with same style like scatter, bubble, time series, heatmaps, treemaps, bar charts etc..
* It supports various R objects.
* It supports Highstocks Charts, Choloropleths.
* It do have piping style which is loved by all R users and programmer.
* A large variety of themes with awesome looks.

Lets get down to the business and create some visulaization with Highcharter following the features mentioned above:

## Creating some Basic Charts Using hchart function

hchart is a generic function which take an object and return an highcharter object. There are functions whose behaviour are similar with the functions of ggplot2 package like :

* hchart works like ggplot2's `qplot`.
* hc_add_series works like ggplot2's `geom_S`.
* hcaes works like ggplot2's `aes`.

Lets choose a dataset, I am going to take Pokemon dataset also provided in Highcharter package.
Have a glimpse of the datset.

```{r, message=FALSE, warning=FALSE}

glimpse(pokemon)

```

Lets plot a bar chart.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "bar", hcaes(x = type_1, y = n))

```


So you got a bar chart with repect of of type 1 category of the pokemons.

Suppose you want a column chart instead then the only variable that you need to change is the type to column.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "column", hcaes(x = type_1, y = n))


```

### Treemap

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "treemap", hcaes(x = type_1, value = n, color = n))


```

We can also use `hc_add_series` to plot the chart. It is used for adding and removing series from highchart object.

### Scatterplot
```{r, message=FALSE, warning=FALSE}

highchart()%>%
  hc_add_series(pokemon, "scatter", hcaes(x = height, y = weight))

```

The main difference in ggplot2's geom_ functions and hc_add_series is that we need to add data and aesthetics explicitly in every function while in ggplot2 one can add data and aesthetics in a layer and then can further add more geoms which can work on same data and aesthetics.


A accurate example is given below using diamond dataset in ggplot2 package.

```{r message=FALSE, warning=FALSE}

data(diamonds, package = "ggplot2")

set.seed(123)
data <- sample_n(diamonds, 300)

modlss <- loess(price ~ carat, data = data)
fit <- arrange(augment(modlss), carat)

highchart() %>% 
  hc_add_series(data, type = "scatter",
                hcaes(x = carat, y = price, size = depth, group = cut)) %>%
  hc_add_series(fit, type = "line", hcaes(x = carat, y = .fitted),
                name = "Fit", id = "fit") %>% 
  hc_add_series(fit, type = "arearange",
                hcaes(x = carat, low = .fitted - 2*.se.fit,
                      high = .fitted + 2*.se.fit),
                linkedTo = "fit")

```

As given in the example the graph is plotted using three series comprising `Scatterplot`, `line`, and `Arearange` 


Lets replicate a highcharts developement into R.



