---
title: "Data Visualization with Highcharter in R"
author: "Nishant"
date: "June 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(highcharter)
library(ggplot2)
library(dplyr)
library(broom)
library(quantmod)
library(tidyverse)
library(lubridate)
data("pokemon")
```


# Data Visualization with Highcharter in R

## Learn how to use Highcharter to create visualization that creates a buzz

I think data visualization is the best technique to show any descriptive and analytics reporting on a chunk of data.
I am a type of person who loves data visualization, you can show the whole story in single screen well that also depends on the data complexity. 
If you are reading this tutorial then I think you must be aware of `Ggplot2` package in R which is used to generate some awesome charts for analysis but somehow lacks dynamic properties.


Coming Back to Highcharter, So it is a `R` wrapper of **HighCharts** javascript library and its module.

The main features of this package are:

* You can create various charts with same style like scatter, bubble, time series, heatmaps, treemaps, bar charts etc..
* It supports various R objects.
* It supports Highstocks Charts, Choloropleths.
* It do have piping style which is loved by all R users and programmer.
* A large variety of themes with awesome looks.

Lets get down to the business and create some visulaization with Highcharter following the features mentioned above:

## Creating some Basic Charts Using hchart function

hchart is a generic function which take an object and return an highcharter object. There are functions whose behaviour are similar with the functions of ggplot2 package like :

* hchart works like ggplot2's `qplot`.
* hc_add_series works like ggplot2's `geom_S`.
* hcaes works like ggplot2's `aes`.

Lets choose a dataset, I am going to take Pokemon dataset also provided in Highcharter package.
Have a glimpse of the datset.

```{r, message=FALSE, warning=FALSE}

glimpse(pokemon)

```

Lets plot a bar chart.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "bar", hcaes(x = type_1, y = n))

```


So you got a bar chart with repect of of type 1 category of the pokemons.

Suppose you want a column chart instead then the only variable that you need to change is the type to column.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "column", hcaes(x = type_1, y = n))


```

### Treemap

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "treemap", hcaes(x = type_1, value = n, color = n))


```

We can also use `hc_add_series` to plot the chart. It is used for adding and removing series from highchart object.

### Scatterplot
```{r, message=FALSE, warning=FALSE}

highchart()%>%
  hc_add_series(pokemon, "scatter", hcaes(x = height, y = weight))

```

The main difference in ggplot2's geom_ functions and hc_add_series is that we need to add data and aesthetics explicitly in every function while in ggplot2 one can add data and aesthetics in a layer and then can further add more geoms which can work on same data and aesthetics.


A accurate example is given below using diamond dataset in ggplot2 package.

```{r message=FALSE, warning=FALSE}

data(diamonds, package = "ggplot2")

set.seed(123)
data <- sample_n(diamonds, 300)

modlss <- loess(price ~ carat, data = data)
fit <- arrange(augment(modlss), carat)

highchart() %>% 
  hc_add_series(data, type = "scatter",
                hcaes(x = carat, y = price, size = depth, group = cut)) %>%
  hc_add_series(fit, type = "line", hcaes(x = carat, y = .fitted),
                name = "Fit", id = "fit") %>% 
  hc_add_series(fit, type = "arearange",
                hcaes(x = carat, low = .fitted - 2*.se.fit,
                      high = .fitted + 2*.se.fit),
                linkedTo = "fit")

```

As given in the example the graph is plotted using three series comprising `Scatterplot`, `line`, and `Arearange` 


Lets replicate a highcharts javasript developement into R using hc_add_series.


>{
>  chart: {
>      type: "area"
>  },
>  title: {
>      text: "Historic and Estimated Worldwide Population Distribution >by Region"
>  },
>  subtitle: {
>      text: "Source: Wikipedia.org"
>  },
>  xAxis: {
>      categories: ["1750", "1800", "1850", "1900", "1950", "1999", >"2050"],
>      tickmarkPlacement: "on",
>      title: {
>          enabled: false
>      }
>  },
>  yAxis: {
>      title: {
>          text: "Percent"
>      }
>  },
>  tooltip: {
>      pointFormat: "<span >style="color:{series.color}">{series.name}</span>:
>      <b>{point.percentage:.1f}%</b> ({point.y:,.0f} millions)<br/>",
>      shared: true
>  },
>  plotOptions: {
>      area: {
>          stacking: "percent",
>          lineColor: "#ffffff",
>          lineWidth: 1,
>          marker: {
>              lineWidth: 1,
>              lineColor: "#ffffff"
>          }
>      }
>  },
>  series: [{
>      name: "Asia",
>      data: [502, 635, 809, 947, 1402, 3634, 5268]
>  }, {
>      name: "Africa",
>      data: [106, 107, 111, 133, 221, 767, 1766]
>  }, {
>      name: "Europe",
>      data: [163, 203, 276, 408, 547, 729, 628]
>  }, {
>      name: "America",
>      data: [18, 31, 54, 156, 339, 818, 1201]
>  }, {
>      name: "Oceania",
>     data: [2, 2, 2, 6, 13, 30, 46]
>  }]
>}

```{r, message=FALSE, warning=FALSE}

highchart() %>% 
  hc_chart(type = "area") %>% 
  hc_title(text = "Historic and Estimated Worldwide Population Distribution by Region") %>% 
  hc_subtitle(text = "Source: Wikipedia.org") %>% 
  hc_xAxis(categories = c("1750", "1800", "1850", "1900", "1950", "1999", "2050"),
           tickmarkPlacement = "on",
           title = list(enabled = FALSE)) %>% 
  hc_yAxis(title = list(text = "Percent")) %>% 
  hc_tooltip(pointFormat = "<span style=\"color:{series.color}\">{series.name}</span>:
             <b>{point.percentage:.1f}%</b> ({point.y:,.0f} millions)<br/>",
             shared = TRUE) %>% 
  hc_plotOptions(area = list(
     stacking = "percent",
     lineColor = "#ffffff",
     lineWidth = 1,
     marker = list(
       lineWidth = 1,
       lineColor = "#ffffff"
       ))
     ) %>% 
  hc_add_series(name = "Asia", data = c(502, 635, 809, 947, 1402, 3634, 5268)) %>% 
  hc_add_series(name = "Africa", data = c(106, 107, 111, 133, 221, 767, 1766)) %>%
  hc_add_series(name = "Europe", data = c(163, 203, 276, 408, 547, 729, 628)) %>% 
  hc_add_series(name = "America", data = c(18, 31, 54, 156, 339, 818, 1201)) %>% 
  hc_add_series(name = "Oceania", data = c(2, 2, 2, 6, 13, 30, 46)) 


```

While comparing you can see each of the block of javascript code convereted into R functions pipelined to each other.
We cab see few arguments like hctooltip's pointFormat in the format used as same in Javascript code you can look into details in [this](http://api.highcharts.com/highcharts#tooltip).


## Highstocks

Highstocks are charts for Financial and time seies analysis, it works well with `quamtmod` library and it's easy chart symbols and then you can add more series using `hc_add_series`.

```{r, message=FALSE, warning=FALSE}

x <- getSymbols("GOOG", auto.assign = FALSE)

hchart(x)

```

As in the chart we don't need to add additional code, hchart accomodated with xts object very efficiently providing a dynamic snapshot of the data.
You can use the zoom functionality to drill down the data in smaller chunks for better analysis.

Lets use hc_add_series.

```{r, message=FALSE, warning=FALSE}

y <- getSymbols("AMZN", auto.assign = FALSE)

highchart(type = "stock") %>% 
  hc_add_series(x) %>% 
  hc_add_series(y, type = "ohlc")

```

As you can see the visulalization adjusting the large volume of data very efficiently.

Do try the stock with different xts objects.

## Highmaps

The easiest way to chart a map with highcharter is using `hcmap` function. Select a url from the highmaps [collection](https://code.highcharts.com/mapdata/) and use the url as a map in hcmap function. This will download the map and create a object using the info as a mapdata argument.

Lets plot India's map.

```{r, message=FALSE, warning=FALSE}

hcmap("https://code.highcharts.com/mapdata/countries/in/in-all.js")%>%
  hc_title(text = "India")

```

Well that's a plain map what about converting it into choropleths.

Every map data downloaded from highcharts map collection have keys to join data. There are 2 functions to help to know what are the regions coded to know how to join the map and data:

* download_map_data: Download the geojson data from the highcharts collection.
* get_data_from_map: Get the properties for each region in the map, as the keys from the map data.

```{r, message=FALSE, warning=FALSE}

mapdata <- get_data_from_map(download_map_data("https://code.highcharts.com/mapdata/countries/in/in-all.js"))

glimpse(mapdata)

#population state wise
pop = as.data.frame(c(84673556, 1382611, 31169272, 103804637, 1055450, 25540196, 342853, 242911, 18980000, 1457723, 60383628, 25353081, 6864602, 
12548926, 32966238, 61130704, 33387677, 64429, 72597565, 112372972, 2721756, 2964007, 1091014, 1980602, 41947358, 1244464, 
27704236, 68621012, 607688, 72138958, 3671032, 207281477, 10116752,91347736))

state= mapdata%>%
  select(`hc-a2`)%>%
  arrange(`hc-a2`)
 
State_pop = as.data.frame(c(state, pop))
names(State_pop)= c("State", "Population")

hcmap("https://code.highcharts.com/mapdata/countries/in/in-all.js", data = State_pop, value = "Population",
      joinBy = c("hc-a2", "State"), name = "Fake data",
      dataLabels = list(enabled = TRUE, format = '{point.name}'),
      borderColor = "#FAFAFA", borderWidth = 0.1,
      tooltip = list(valueDecimals = 0)) 

```

Do some experiments with hc_add_series in the maps and choropleths. For some details and examples visit [here](http://jkunst.com/highcharter/highmaps.html).

## PlugIns 

Now lets try some plug ins provided by highcharter like grouping, drilldowns, downloading, printing data, and some awesome themes.

Lets group data of mpg dataset, here for better visiualisation we created a list which categorises data according to the manufacturer.

```{r message=FALSE, warning=FALSE}

data(mpg, package = "ggplot2")

mpgg <- mpg %>% 
  filter(class %in% c("suv", "compact", "midsize")) %>% 
  group_by(class, manufacturer) %>% 
  summarize(count = n())

categories_grouped <- mpgg %>% 
  group_by(name = class) %>% 
  do(categories = .$manufacturer) %>% 
  list_parse()

highchart() %>% 
  hc_xAxis(categories = categories_grouped) %>% 
  hc_add_series(data = mpgg, type = "bar", hcaes(y = count, color = manufacturer),
                showInLegend = FALSE)

```

Lets create a chart which drill downs to another chart for better analysis.
You can understand the code more efficiently if you understand lists in R.

```{r message=FALSE, warning=FALSE}

df <- data_frame(
  name = c("Animals", "Fruits", "Cars"),
  y = c(5, 2, 4),
  drilldown = tolower(name)
)


ds <- list_parse(df)
names(ds) <- NULL


hc <- highchart() %>% 
  hc_chart(type = "column") %>% 
  hc_title(text = "Basic drilldown") %>% 
  hc_xAxis(type = "category") %>% 
  hc_legend(enabled = FALSE) %>% 
  hc_plotOptions(
    series = list(
      boderWidth = 0,
      dataLabels = list(enabled = TRUE)
    )
  ) %>% 
  hc_add_series(
    name = "Things",
    colorByPoint = TRUE,
    data = ds
  )

dfan <- data_frame(
  name = c("Cats", "Dogs", "Cows", "Sheep", "Pigs"),
  value = c(4, 3, 1, 2, 1)
)

dffru <- data_frame(
  name = c("Apple", "Organes"),
  value = c(4, 2)
)

dfcar <- data_frame(
  name = c("Toyota", "Opel", "Volkswage"),
  value = c(4, 2, 2) 
)

second_el_to_numeric <- function(ls){
  
  map(ls, function(x){
    x[[2]] <- as.numeric(x[[2]])
    x
  })
  
}

dsan <- second_el_to_numeric(list_parse2(dfan))

dsfru <- second_el_to_numeric(list_parse2(dffru))

dscar <- second_el_to_numeric(list_parse2(dfcar))

hc %>% 
  hc_drilldown(
    allowPointDrilldown = TRUE,
    series = list(
      list(
        id = "animals",
        data = dsan
      ),
      list(
        id = "fruits",
        data = dsfru
      ),
      list(
        id = "cars",
        data = dscar
      )
    )
  )

```

Lets apply Drill down on other charts.

```{r message=FALSE, warning=FALSE}

tm <- pokemon %>% 
  mutate(type_2 = ifelse(is.na(type_2), paste("only", type_1), type_2),
         type_1 = type_1) %>% 
  group_by(type_1, type_2) %>%
  summarise(n = n()) %>% 
  ungroup() %>% 
  treemap::treemap(index = c("type_1", "type_2"),
                   vSize = "n", vColor = "type_1")

tm$tm <- tm$tm %>%
  tbl_df() %>% 
  left_join(pokemon %>% select(type_1, type_2, color_f) %>% distinct(), by = c("type_1", "type_2")) %>%
  left_join(pokemon %>% select(type_1, color_1) %>% distinct(), by = c("type_1")) %>% 
  mutate(type_1 = paste0("Main ", type_1),
         color = ifelse(is.na(color_f), color_1, color_f))

highchart() %>% 
  hc_add_series_treemap(tm, allowDrillToNode = TRUE,
                        layoutAlgorithm = "squarified")

```

Lets add functionality of exporting data.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "bar", hcaes(x = type_1, y = n, color = type_1))%>%
  hc_exporting(enabled = TRUE)

```

At last we add some themes.

```{r message=FALSE, warning=FALSE}

pokemon%>%
  count(type_1)%>%
  arrange(n)%>%
  hchart(type = "bar", hcaes(x = type_1, y = n, color = type_1))%>%
  hc_exporting(enabled = TRUE)%>%
hc_add_theme(hc_theme_chalk())

```
You can learn a lot from [here](http://jkunst.com/highcharter/themes.html), Its a complete library for Highcharter.
I am also sharing one of my favourite graph created on weather dataset, here one of the intersting argument `polar` is used by setting the argument to TRUE the whole story is transformed.

```{r message=FALSE, warning=FALSE}

data("weather")

x <- c("Min", "Mean", "Max")
y <- sprintf("{point.%s}", c("min_temperaturec", "mean_temperaturec", "max_temperaturec"))
tltip <- tooltip_table(x, y)

hchart(weather, type = "columnrange",
       hcaes(x = date, low = min_temperaturec, high = max_temperaturec,
             color = mean_temperaturec)) %>% 
  hc_chart(polar = TRUE) %>%
  hc_yAxis( max = 30, min = -10, labels = list(format = "{value} C"),
            showFirstLabel = FALSE) %>% 
  hc_xAxis(
    title = list(text = ""), gridLineWidth = 0.5,
    labels = list(format = "{value: %b}")) %>% 
  hc_tooltip(useHTML = TRUE, pointFormat = tltip,
             headerFormat = as.character(tags$small("{point.x:%d %B, %Y}")))

```

